Analysis of celiac disease

### 2018-09-21 4-results-simus.Rmd
change names

---
title: "Results"
author: "Florian Privé"
date: "November 17, 2017"
output: 
  html_document:
    code_folding: hide
---

@@

```{r}
library(tidyverse)
# devtools::install_github("privefl/bigsnpr")
library(bigsnpr)
# devtools::install_github("privefl/paper2-PRS/pkg.paper.PRS")
library(pkg.paper.PRS)  ## See code source for functions' code.
options(nboot = 1e3)
```

@@

```{r}
compare_logit <- function(results, with_method) {

  results <- filter(
    results, method %in% c("logit-simple", with_method), par.h2 == 0.8)

@@

```{r, out.width="90%", fig.width = 14}
compare_logit(results1, with_method = "T-Trees")

@@

## Results with logit-triple & model fancy

@@

### Results with logit-triple

```{r, out.width="90%", fig.width = 14}
compare_logit(results2, with_method = "logit-triple")

@@

```{r}
results2 %>%
  filter(grepl("logit", method)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_one(h2 = 0.5) +
  facet_grid(par.model ~ par.dist)

@@

### Results with model fancy

```{r}
results2 %>%
  filter(par.model == "fancy", method %in% c("logit-simple", "PRS-max")) %>%
  filter(thr.r2 == 0.2 | is.na(thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_all()

@@

```{r}
results2 %>%
  filter(par.model == "fancy", grepl("PRS", method)) %>%
  filter(thr.r2 == 0.2 | is.na(thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_all()

@@

## Results with only logit-simple & PRSs

@@

```{r}
results2.all.r2 <- results2 %>%
  filter(par.model == "simple", method %in% c("logit-simple", "PRS-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n())

@@

```{r}
results2 %>%
  filter(par.model == "simple", method %in% c("logit-simple", "PRS-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AltSens_mean = mean(AltSens), AltSens_boot = boot(AltSens), N = n()) %>%
  myggplot(aes(par.causal, AltSens_mean, fill = method, color = method)) + 
  geom_col(position = position_dodge(), alpha = 0.3) +
  geom_errorbar(aes(ymin = AltSens_mean - 2 * AltSens_boot,
                    ymax = AltSens_mean + 2 * AltSens_boot),
                position = position_dodge(width = 0.9), color = "black", width = 0.2) + 
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Causal SNPs (number and location)", y = "Mean of 100 AltSens",
       fill = "Method", color = "Method") +
  scale_fill_manual(values = methods.color) +
  scale_color_manual(values = methods.color) +
  facet_grid(par.dist ~ par.h2)

@@

```{r}
results3 <- results2 %>%
  filter(thr.r2 == 0.2 | is.na(thr.r2)) %>%
  filter(par.model == "simple", method != "logit-triple")
```

```{r, fig.asp=0.7, out.width="95%", fig.width=16}
results3 %>%
  filter(method %in% c("logit-simple", "PRS-max")) %>%
  mutate(Parameters  = interaction(par.h2, par.dist, method, sep = " | ")) %>%
  myggplot(aes(AUC, pAUC, color = Parameters)) + 
  geom_point(size = 0.6, alpha = 0.5) +
  geom_smooth(aes(linetype = Parameters), method = "loess", se = FALSE, size = 2) +
  theme(legend.key.width = unit(3.5, "line"), 
        legend.direction = "horizontal") + 
  labs(y = "Partial AUC (specificity: [0.9, 1])") +
  theme(legend.position = "top", axis.title = element_text(size = rel(2.2))) +

@@

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(method %in% c("logit-simple", "PRS-max"), par.model == "simple") %>%
  barplot_causal_one(h2 = 0.8)

@@

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(grepl("PRS", method)) %>%
  barplot_causal_one(h2 = 0.8)

@@

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(method %in% c("logit-simple", "PRS-max")) %>%
  barplot_causal_one(h2 = 0.5)

@@

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(grepl("PRS", method)) %>%
  barplot_causal_one(h2 = 0.5)

@@

```{r}
results4 %>%
  filter(method %in% c("logit-simple", "PRS-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_all() + 
  geom_col(aes(par.causal, AUC_mean), data = results2.all.r2, 
           position = position_dodge(), color = "black", alpha = 0)

@@

results6 <- results2 %>%
  filter(par.model == "simple", par.causal == "300 in all") %>%
  bind_rows(results5) %>%
  filter(method %in% c("logit-simple", "PRS-max")) %>%
  mutate(method = c_method_r2(method, thr.r2))
```

@@

```{r, fig.asp=0.65, out.width="95%", fig.width=16}
list.files("results7", full.names = TRUE) %>%
  read_format_results(corr) %>%
  compare_logit("biglasso")

>>>>

---
title: "Results"
author: "Florian Privé"
date: "September 21, 2018"
output: 
  html_document:
    code_folding: hide
---

@@

```{r}
library(tidyverse)
# devtools::install_github("privefl/bigsnpr")
library(bigsnpr)
# devtools::install_github("privefl/paper2-PRS/pkg.paper.PRS")
library(pkg.paper.PRS)  ## See code source for functions' code.
options(nboot = 1e5)
```

@@

```{r}
compare_PLR <- function(results, with_method) {

  results <- filter(
    results, method %in% c("PLR", with_method), par.h2 == 0.8)

@@

```{r, out.width="90%", fig.width = 14}
compare_PLR(results1, with_method = "T-Trees")

@@

## Results with PLR3 & model COMP

@@

### Results with PLR3

```{r, out.width="90%", fig.width = 14}
compare_PLR(results2, with_method = "PLR3")

@@

```{r}
results2 %>%
  filter(grepl("PLR", method)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_one(h2 = 0.5) +
  facet_grid(par.model ~ par.dist)

@@

### Results with model COMP

```{r}
results2 %>%
  filter(par.model == "COMP", method %in% c("PLR", "C+T-max")) %>%
  filter(thr.r2 == 0.2 | is.na(thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_all()

@@

```{r}
results2 %>%
  filter(par.model == "COMP", grepl("C\\+T", method)) %>%
  filter(thr.r2 == 0.2 | is.na(thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_all()

@@

## Results with only PLR & C+T

@@

```{r}
results2.all.r2 <- results2 %>%
  filter(par.model == "ADD", method %in% c("PLR", "C+T-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n())

@@

```{r}
results2 %>%
  filter(par.model == "ADD", method %in% c("PLR", "C+T-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AltSens_mean = mean(AltSens), AltSens_boot = boot(AltSens), N = n()) %>%
  myggplot(aes(par.causal, AltSens_mean, fill = method, color = method)) + 
  geom_col(position = position_dodge(), alpha = 0.3) +
  geom_errorbar(aes(ymin = AltSens_mean - 2 * AltSens_boot,
                    ymax = AltSens_mean + 2 * AltSens_boot),
                position = position_dodge(width = 0.9), color = "black", width = 0.2) + 
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Causal SNPs (number and location)", y = "Mean of 100 AltSens",
       fill = "Method", color = "Method") +
  scale_fill_manual(values = methods.color) +
  scale_color_manual(values = methods.color) +
  facet_grid(par.dist ~ par.h2)

@@

```{r}
results3 <- results2 %>%
  filter(thr.r2 == 0.2 | is.na(thr.r2)) %>%
  filter(par.model == "ADD", method != "PLR3")
```

```{r, fig.asp=0.7, out.width="95%", fig.width=16}
results3 %>%
  filter(method %in% c("PLR", "C+T-max")) %>%
  mutate(Parameters  = interaction(par.h2, par.dist, method, sep = " | ")) %>%
  myggplot(aes(AUC, pAUC, color = Parameters)) + 
  geom_point(size = 0.6, alpha = 0.5) +
  geom_smooth(aes(linetype = Parameters), method = "loess", se = FALSE, size = 2) +
  theme(legend.key.width = unit(3.5, "line"), 
        legend.direction = "horizontal") + 
  labs(y = "Partial AUC (specificity: [0.9, 1])") +
  theme(legend.position = "top", axis.title = element_text(size = rel(2.2))) +

@@

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(method %in% c("PLR", "C+T-max"), par.model == "ADD") %>%
  barplot_causal_one(h2 = 0.8)

@@

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(grepl("C\\+T", method)) %>%
  barplot_causal_one(h2 = 0.8)

@@

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(method %in% c("PLR", "C+T-max")) %>%
  barplot_causal_one(h2 = 0.5)

@@

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(grepl("C\\+T", method)) %>%
  barplot_causal_one(h2 = 0.5)

@@

```{r}
results4 %>%
  filter(method %in% c("PLR", "C+T-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_all() + 
  geom_col(aes(par.causal, AUC_mean), data = results2.all.r2, 
           position = position_dodge(), color = "black", alpha = 0)

@@

results6 <- results2 %>%
  filter(par.model == "ADD", par.causal == "300 in all") %>%
  bind_rows(results5) %>%
  filter(method %in% c("PLR", "C+T-max")) %>%
  mutate(method = c_method_r2(method, thr.r2),
         par.causal = as.character(par.causal))
```

@@

```{r, fig.asp=0.65, out.width="95%", fig.width=16}
list.files("results7", full.names = TRUE) %>%
  read_format_results(corr) %>%
  compare_PLR("biglasso")

### 2018-09-21 5-results-celiac.Rmd
change names

```{r}
library(tidyverse)
library(bigsnpr)
library(pkg.paper.PRS)
```

@@

### PRS

@@

```{r, cache=TRUE}
# PRS
thrs <- c(0, -log10(5e-8), exp(seq(log(0.1), log(100), length.out = 100)))
lpS <- -predict(gwas.train.gc)
prs <- snp_PRS(G, betas.keep = gwas.train.gc$estim[ind.keep],
               ind.test = ind.test,
               ind.keep = ind.keep,
               lpS.keep = lpS[ind.keep], 
               thr.list = thrs)
nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr))
aucs <- apply(prs, 2, AUC, target = y[ind.test])
summary(aucs)
nb.pred[which.max(aucs)]
```

```{r}
pAUC <- pkg.paper.PRS:::pAUC

plot_density_AUC <- function(pred, true = y[ind.test]) {
  true_fct <- forcats::fct_recode(as.factor(true), "Control" = "0", "Case" = "1")
  data.frame(Score = pred, Status = true_fct) %>%
  ggplot() +
  geom_density(aes(Score, fill = Status), alpha = 0.4) +
  ggtitle(sprintf("AUC = %s%% | pAUC = %s", 
                  round(100 * AUC(pred, true), 2),
                  round(pAUC(pred, true), 4))) + 
    theme_bigstatsr() +
    theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
}
(p1 <- plot_density_AUC(prs[, which.max(aucs)]))
```

@@

```{r, cache=TRUE}
# Train logit-simple
system.time(
  logit <- big_spLogReg(X = G, y01.train = y[ind.train], 
                        ind.train = ind.train, 
                        covar.train = obj.svd$u[ind.train, ],
                        ncores = NCORES, alphas = 1)
)
# Get K = 10 predictions/individuals
preds1 <- predict(logit, X = G, ind.row = ind.test, 
                  covar.row = obj.svd$u[ind.test, ])
(nb.pred1 <- sum(rowSums(sapply(logit, function(l) l$beta.X != 0)) > 0))
(aucs1 <- AUC(preds1, target = y[ind.test]))
```

@@

### Logistic regression with feature engineering

@@

```{r, fig.asp=1, fig.width=8}
library(plotROC)
scores_tidy <- tibble(
  d = y[ind.test], 
  "PRS-max"      = prs[, which.max(aucs)],
  "logit-simple" = preds1,
  "logit-triple" = preds2
) %>%
  gather(key = "Method", value = "Score", -d)

methods.color <- setNames(scales::hue_pal()(6), 
                          c("logit-simple", "PRS-max", "PRS-stringent",
                            "T-Trees", "logit-triple", "PRS-all"))

@@

```{r}
list.files(path = "results6/", full.names = TRUE) %>%
  map_df(readRDS) %>%
  as_tibble() %>%
  mutate(
    AUC = map2_dbl(pred, pheno, ~bigstatsr::AUC(.x, .y)),
    pAUC = map2_dbl(pred, pheno, ~pAUC(.x, .y)),
    nb.preds = map_int(set, length)
  ) %>%
  group_by(method) %>%
  summarize_at(c("AUC", "pAUC", "nb.preds", "timing"), function(x) {
    glue::glue("{signif(mean(x), 3)} ({signif(boot(x), 3)})")
  }) %>%
  print() %>%
  xtable::xtable()
```

>>>>

```{r}
library(tidyverse)
library(bigsnpr)
library(pkg.paper.PRS)
pAUC <- pkg.paper.PRS:::pAUC
```

@@

### C+T

@@

```{r, cache=TRUE}
# C+T
thrs <- c(0, -log10(5e-8), exp(seq(log(0.1), log(100), length.out = 100)))
lpS <- -predict(gwas.train.gc)
prs <- snp_PRS(G, betas.keep = gwas.train.gc$estim[ind.keep],
               ind.test = ind.test,
               ind.keep = ind.keep,
               lpS.keep = lpS[ind.keep], 
               thr.list = thrs)
nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr))
aucs <- apply(prs, 2, AUC, target = y[ind.test])
summary(aucs)
nb.pred[which.max(aucs)]
```

```{r}
plot_density_AUC <- function(pred, true = y[ind.test]) {
  true_fct <- forcats::fct_recode(as.factor(true), Control = "0", Case = "1")
  data.frame(Score = pred, Status = true_fct) %>%
  ggplot() +
  geom_density(aes(Score, fill = Status), alpha = 0.4) +
  ggtitle(sprintf("AUC = %s%% | pAUC = %s", 
                  round(100 * AUC(pred, true), 2),
                  round(pAUC(pred, true), 4))) + 
    theme_bigstatsr() +
    theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
}
(p1 <- plot_density_AUC(prs[, which.max(aucs)]))
```

@@

```{r, cache=TRUE}
# Train PLR
system.time(
  logit <- big_spLogReg(X = G, y01.train = y[ind.train], 
                        ind.train = ind.train, 
                        covar.train = obj.svd$u[ind.train, ],
                        ncores = NCORES, alphas = 1)
)
# Get K = 10 predictions/individuals
preds1 <- predict(logit, X = G, ind.row = ind.test, 
                  covar.row = obj.svd$u[ind.test, ])
(nb.pred1 <- sum(rowSums(sapply(logit, function(l) l$beta.X != 0)) > 0))
(aucs1 <- AUC(preds1, target = y[ind.test]))
```

@@

### Logistic regression with feature engineering (PLR3)

@@

```{r, fig.asp=1, fig.width=8}
library(plotROC)
scores_tidy <- tibble(
  d = y[ind.test], 
  "C+T-max" = prs[, which.max(aucs)],
  "PLR"     = preds1,
  "PLR3"    = preds2
) %>%
  gather(key = "Method", value = "Score", -d)

methods.color <- setNames(scales::hue_pal()(6), 
                          c("PLR", "C+T-max", "C+T-stringent",
                            "T-Trees", "PLR3", "C+T-all"))

@@

```{r}
list.files(path = "results6/", full.names = TRUE) %>%
  map_df(readRDS) %>%
  as_tibble() %>%
  mutate(
    method = fct_recode(sub("PRS", "C+T", method), 
                        PLR = "logit-simple", PLR3 = "logit-triple"),
    AUC = map2_dbl(pred, pheno, ~bigstatsr::AUC(.x, .y)),
    pAUC = map2_dbl(pred, pheno, ~pAUC(.x, .y)),
    nb.preds = map_int(set, length)
  ) %>%
  group_by(method) %>%
  summarize_at(c("AUC", "pAUC", "nb.preds", "timing"), function(x) {
    glue::glue("{signif(mean(x), 3)} ({signif(boot(x), 3)})")
  }) %>%
  print() %>%
  xtable::xtable()
```
